---
title: "Supplemental Materials for The voice of the parrot: vocal identity across multiple call types in a communally nesting parrot"
output:
  bookdown::pdf_document2:
    toc: yes
editor_options:
  chunk_output_type: console
header-includes:
- \renewcommand{\figurename}{Figure S}
- \makeatletter
- \def\fnum@figure{\figurename\thefigure}
- \makeatother
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
# Set up
library(tidyverse)
library(scales)

# Clean R
rm(list=ls()) 

# Paths
path_pdf_pdfa = '../../RESULTS/04_cross_call_type/scores.pdf'
path_pdf_pdfa_sub_f = '../../RESULTS/04_cross_call_type/scores_sub_f.pdf'
path_pdf_pdfa_sub_m = '../../RESULTS/04_cross_call_type/scores_sub_m.pdf'
path_pdf_pdfa_permuted = '../../RESULTS/04_cross_call_type/scores_permuted.pdf'
path_pdf_year = '../../RESULTS/03_year_comparison/results.pdf'
path_pdf_spectros = '../../RESULTS/figures/example spectrograms.pdf'
```

# Supplemental Methods

## Call types

Call types were manually determined from the spectrogram and audio in Raven Lite [@RavenLite]. We included X types in the main text (see also figure SX):

- contact: tonal vocalisation component and at least three frequency modulation peaks
- tja: tonal vocalisation with a single up-sweep component; sometimes one or two very short amplitude modulated components are present before or after the main component
- trruup: vocalisation that consists of at least three amplitude modulated components at the start, followed by a frequency modulated tonal component
- alarm: loud amplitude modulated tonal vocalisation with at least five clear u-shaped components
- growl: medium loud amplitude modulated vocalisation with at least five components without clear tonality

```{r figure-spectros, include=TRUE, fig.align="center", fig.cap='Example spectrograms of the call types.', fig.height=5, fig.width=7, echo=FALSE}
knitr::include_graphics(path_pdf_spectros)
```

\pagebreak

## Dynamic time warping (DTW)

For tonal calls the fundamental frequency was traced in Luscinia  [@lachlan2007luscinia]. For non-tonal calls only the start and end times were determined manually. Traces were then exported as CSV and imported into R [@R] for further analysis. Traces were smoothed using two steps. The first step filled gaps within a call by connecting the fundamental frequency at both sides of the gap with a straight line. To get rid of small measurement jitter, we used smooth.spline from *stats* with spar = 0.1. 

[ADD DETAILS SPCC, DTW, MFCCCC and SPECAN]

To measure the similarity between any two calls a variety of methods have been used. The most important are 1) dynamic time warping (DTW), which measures measures similarity between the fundamental frequencies of two calls while controlling for differences in speech rhythm \parencite{Bellman1959IRE}; 2) spectrographic cross correlation (SPCC), which measures similarity between the spectrograms of two calls \parencite{clark1987quantitative}; and 3) spectrograpic analysis (SPECAN) which measures a set of variables on the spectrogram. We added a fourth method not commonly used and cross correlated the mel frequency cepstral coefficients (MFCCCC). 

## Statistical analysis

### Model I: individual and recording signal

We used the dis-similarity matrices from DTW, SPCC, MFCCCC or SPECAN to get a pairwise acoustic distance between all calls. We only included call pairs from the same year. We used a model similar to the Bayesian social relations model used in psychology. In our case the acoustic distance between any two calls was the response variable and was standardised before analysis. We included a global intercept and off-sets for: same vs different individual, same vs different recording, pair of individuals from which the calls came, pair of recordings from which the calls came and each call. The full model structure was as follows:

$$
\begin{aligned}
\text{accoustic distance} & \sim \text{normal}(\mu,\ \sigma)\\
\mu_n & = \bar\alpha + \alpha_\text{same ind[n]} + \alpha_\text{same rec[n]} +\\
& \ \ \ \ \ \ \alpha_\text{ind pair[n]} + \alpha_\text{rec pair[n]} + \alpha_\text{call i[n]} + \alpha_\text{call j[n]}\\
\bar\alpha & \sim \text{normal}(0,\ 0.25)\\
\alpha_\text{same ind} & \sim \text{normal}(0,\ \sigma_\text{same ind})\\
\alpha_\text{same rec} & \sim  \text{normal}(0,\ \sigma_\text{same rec})\\
\alpha_\text{ind pair} & \sim  \text{normal}(0,\ \sigma_\text{ind pair})\\
\alpha_\text{rec pair} & \sim  \text{normal}(0,\ \sigma_\text{rec pair})\\
\alpha_\text{call} & \sim  \text{normal}(0,\ \sigma_\text{call})\\
\sigma_\text{same ind},\ \sigma_\text{same rec},\ \sigma_\text{ind pair},\ \sigma_\text{rec pair} & \sim \text{exponential}(3)\\
\sigma,\ \sigma_\text{call} & \sim \text{exponential}(5)
\end{aligned}
$$

We fitted an un-centred version of the model using the package *cmdstanr* [@cmdstanr] with the No U-turn Sampler in Stan [@gelman2015stan] from R [@R]. We ran 2000 iterations on four chains with adapt_delta = 0.99 and max_treedepth = 15. We used the difference between $\alpha_\text{same ind [same]}$ and $\alpha_\text{same ind [different]}$ as measure of individual signal and the difference between $\alpha_\text{same rec [same]}$ and $\alpha_\text{same rec [different]}$ as measure of recording signal. The other off-sets were included to control for unbalanced data and non-independence (e.g., some individuals might sound very much alike and we included multiple pairs of calls from the same two individuals). 

### Model II: decay over time within recording

To test how stable the individual signature is within a recording we subsetted the pairwise dis-similarities to only include comparisons between calls from the same individual and recording. We included log10 of the time (in minutes) between the calls. We then used a multi-level model with a global intercept, a slope for the time-effect, off-sets for the intercept pair of individuals from which the calls came, pair of recordings from which the calls came and each call. We also included off-sets on the slope for individual. The full model structure was as follows:

$$
\begin{aligned}
\text{accoustic distance} & \sim \text{normal}(\mu,\ \sigma)\\
\mu_n & = \bar\alpha + \alpha_\text{ind[n]} + \alpha_\text{rec[n]} + \alpha_\text{call i[n]} + \alpha_\text{call j[n]}\\
& \ \ \ \ \ \ (\bar\beta + \beta_\text{ind[n]} + \beta_\text{rec[n]}) * \text{time}_n \\
\bar\alpha & \sim \text{normal}(0,\ 0.5)\\
\alpha_\text{ind} & \sim \text{normal}(0,\ \sigma_\text{ind})\\
\alpha_\text{rec} & \sim  \text{normal}(0,\ \sigma_\text{rec})\\
\alpha_\text{call} & \sim  \text{normal}(0,\ \sigma_\text{call})\\
\bar\beta & \sim \text{normal}(0,\ 0.3)\\
\beta_\text{ind} & \sim \text{normal}(0,\ \xi_\text{ind})\\
\beta_\text{rec} & \sim  \text{normal}(0,\ \xi_\text{rec})\\
\sigma,\ \sigma_\text{ind},\ \sigma_\text{rec},\ \xi_\text{ind},\ \xi_\text{rec} & \sim \text{exponential}(2)\\
\sigma_\text{call} & \sim \text{exponential}(3)
\end{aligned}
$$

The model was fitted in the same way as model I.

### Model III: decay over days

To test how stable the individual signature is across days we subsetted the pairwise dis-similarities to only include comparisons between calls from the same individual and year, but from different recordings. We also included the difference in days between the calls. We then used a multi-level model with a global intercept, a slope for the time-effect, off-sets for the intercept for individual pair, recording and call, and off-sets on the slope for individual. The full model structure was as follows:

$$
\begin{aligned}
\text{accoustic distance} & \sim \text{normal}(\mu,\ \sigma)\\
\mu_n & = \bar\alpha + \alpha_\text{ind pair[n]} + \alpha_\text{rec pair[n]} + \alpha_\text{call i[n]} + \alpha_\text{call j[n]}\\
& \ \ \ \ \ \ (\bar\beta + \beta_\text{ind[n]})* \text{time}_n \\
\bar\alpha & \sim \text{normal}(0,\ 0.5)\\
\alpha_\text{ind pair} & \sim \text{normal}(0,\ \sigma_\text{ind pair})\\
\alpha_\text{rec pair} & \sim  \text{normal}(0,\ \sigma_\text{rec pair})\\
\alpha_\text{call} & \sim  \text{normal}(0,\ \sigma_\text{call})\\
\bar\beta & \sim \text{normal}(0,\ 0.5)\\
\beta_\text{ind} & \sim \text{normal}(0,\ \xi_\text{ind})\\
\sigma,\ \sigma_\text{ind pair},\ \sigma_\text{rec pair},\ \sigma_\text{call},\ \xi_\text{ind} & \sim \text{exponential}(2)
\end{aligned}
$$

The model was fitted in the same way as model I.

### Model IV: difference between years

To test how stable the individual signature is across years we subsetted the pairwise dis-similarities to only include comparisons between calls from the same individual, but from different recordings. We also included whether or not the recordings came from the same year. We then used a multi-level model with a global intercept, off-sets for: same vs different year, the individual, recording pair and call. The full model structure was as follows:

$$
\begin{aligned}
\text{accoustic distance} & \sim \text{normal}(\mu,\ \sigma)\\
\mu_n & = \bar\alpha + \alpha_\text{same year[n]} + \alpha_\text{ind[n]} +\\
& \ \ \ \ \ \ \alpha_\text{rec pair[n]} + \alpha_\text{call i[n]} + \alpha_\text{call j[n]}\\
\bar\alpha & \sim \text{normal}(0,\ 0.25)\\
\alpha_\text{same year} & \sim \text{normal}(0,\ \sigma_\text{same year})\\
\alpha_\text{ind} & \sim  \text{normal}(0,\ \sigma_\text{ind})\\
\alpha_\text{rec pair} & \sim  \text{normal}(0,\ \sigma_\text{rec pair})\\
\alpha_\text{call} & \sim  \text{normal}(0,\ \sigma_\text{call})\\
\sigma,\ \sigma_\text{same year},\ \sigma_\text{ind},\ \sigma_\text{rec pair},\ \sigma_\text{call} & \sim \text{exponential}(3)\\
\end{aligned}
$$

The model was fitted in the same way as model I.

### pDFA

To test if there are features that contain an individual signal across call types we used mel frequency cepstral coefficients (MFCC). More specifically we used the function *melfcc* from the package *warbleR* [@warbleR] with the settings wintime = 512/44100, hoptime = 50/44100 and minfreq = 300 to retrieve the first ten cepstral coefficient traces. For each trace we saved the mean and standard deviation. These 20 measures per call were then used for the pDFA.

We wrote a custom R script to run the pDFA. We ran 100 iterations with the following steps:

- for each individual we randomly selected recordings until the set of recordings contained enough calls for testing (N_test)
- we then selected the other recordings for training 
- if there were not enough recordings to fullfill both N_test and N_train, we did not include this individual
- for the individuals with enough data, we randomly selected N_test calls from the testing recordings and N_train calls from the training recordings
- we then used the standardised MFCC values to train an LDA (using the function *lda* from the package *MASS* [@MASS])
- this LDA was then used to predict the IDs of the test data, the resulting proportion of correct classification was the trained score
- we repeated the previous two steps with both training and testing IDs randomised within the set and retrieved the random score
- for each iteration the difference between the trained and random score was reported

\pagebreak

# Supplemental Results

## Model I: results other methods



## Model IV: difference between years

There was no effect of year for any of the methods or call types (see Figure SX). 

```{r figure-year, include=TRUE, fig.align="center", fig.cap='Posterior distributions for the for the Bayesian multilevel model with difference between recordings from the same vs different years. Greater values  mean more within year similarity. Models were run separately for the main call types (columns) and methods (rows; DTW - dynamic time warping, SPCC - spectrographic cross correlation, MF4C - mel frequency cepstral coefficient cross correlation). N represents the number of calls included in the model.', fig.height=5, fig.width=7, echo=FALSE}
knitr::include_graphics(path_pdf_year)
```

\pagebreak

## pDFA

For the results presented in the paper there was no overlap with zero for any of the pDFAs. For full distributions and sample sizes see Figure SX.

```{r figure-pdfa, include=TRUE, fig.align="center", fig.cap='pDFA results. Densities of the differences between the trained and random DFAs. N is number of individals for which there was enough data available.', fig.height=5, fig.width=7, echo=FALSE}
knitr::include_graphics(path_pdf_pdfa)
```

\pagebreak

When sub-setting the data to only include females from Promenade Passeig de Lluís Companys the results for growly to tonal fell towards chance-level, but all other results remained stable with overlap with zero less than 2%. For full distributions and sample sizes see Figure SX.

```{r figure-pdfa-sub-f, include=TRUE, fig.align="center", fig.cap='pDFA results for females from Promenade Passeig de Lluís Companys. Densities of the differences between the trained and random DFAs. N is number of individals for which there was enough data available.', fig.height=5, fig.width=7, echo=FALSE}
knitr::include_graphics(path_pdf_pdfa_sub_f)
```

\pagebreak

When sub-setting the data to only include males from Promenade Passeig de Lluís Companys the results for both cross-call type models fell towards chance-level. For full distributions and sample sizes see Figure SX.

```{r figure-pdfa-sub-m, include=TRUE, fig.align="center", fig.cap='pDFA results for males from Promenade Passeig de Lluís Companys. Densities of the differences between the trained and random DFAs. N is number of individals for which there was enough data available.', fig.height=5, fig.width=7, echo=FALSE}
knitr::include_graphics(path_pdf_pdfa_sub_m)
```

\pagebreak

When randomising within Promenade Passeig de Lluís Companys and Parc de la Ciutadella the results for the cross call type models remained stable but overlap with zero increased to 6-7%. For full distributions and sample sizes see Figure SX.

```{r figure-pdfa-perm, include=TRUE, fig.align="center", fig.cap='pDFA results with randomisation withing Promenade Passeig de Lluís Companys and Parc de la Ciutadella only. Densities of the differences between the trained and random DFAs. N is number of individals for which there was enough data available.', fig.height=5, fig.width=7, echo=FALSE}
knitr::include_graphics(path_pdf_pdfa_permuted)
```

\pagebreak

# References
